---
title: "SAL 602 Project"
author: "Andrew Fish"
date: "2024-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(shiny)
library(tidyverse)
library(Lahman)
library(baseballr)
library(readr)
library(ggrepel)
library(plotly)
```

The following sections of code is the data scraping and data manipulation of the statcast data used in this project.

```{r}
masterID <- read_csv("C:/Users/Andrew Fish/Documents/Syracuse/SAL 602/Module 7/masterid.csv")
```

```{r}
##scrapes a month with defined start date so only getting regular season data
scrapeStatcastFirstmonth <- function(year, month, start_date, playerID) {
  if (month == 4) {
    scrape1 <- scrape_statcast_savant_batter(start_date,
                                             paste(as.character(year), "-0", as.character(month), "-07", sep = ""),
                                             batterid = playerID)
    
    scrape2 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-08", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-14", sep = ""),
                                             batterid = playerID)
    
    scrape3 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-15", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-21", sep = ""),
                                             batterid = playerID)
    
    scrape4 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-22", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-28", sep = ""),
                                             batterid = playerID)
    
    scrape5 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-29", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-30", sep = ""),
                                             batterid = playerID)
    
    statcastScrape <- rbind(scrape1, scrape2, scrape3, scrape4, scrape5)
  }
  if (month == 3) {
    start_date %>% strsplit("-") -> split_date
    day <- as.numeric(split_date[[1]][3])
    if (30 - day < 7) {
      statcastScrape <- scrape_statcast_savant_batter(start_date,
                                                      paste(as.character(year), "-0", as.character(month), "-30", sep = ""),
                                                      batterid = playerID)
    }
    else if ((30 - day <14) & (30 - day >= 7)) {
      scrape1 <- scrape_statcast_savant_batter(start_date,
                                               paste(as.character(year), "-0", as.character(month), "-23", sep = ""),
                                               batterid = playerID)
      
      scrape2 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-24", sep = ""),
                                               paste(as.character(year), "-0", as.character(month), "-30", sep = ""),
                                               batterid = playerID)
      statcastScrape <- rbind(scrape1, scrape2)
    }
  }
  return(statcastScrape)
}
```

```{r}
##scrapes month with end date so only getting regular season data
scrapeStatcastLastmonth <- function(year, month, end_date, playerID) {
  if (month == 9) {
    scrape1 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-01", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-07", sep = ""),
                                             batterid = playerID)
    scrape2 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-08", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-14", sep = ""),
                                             batterid = playerID)
    scrape3 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-15", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-21", sep = ""),
                                             batterid = playerID)
    scrape4 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-22", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-28", sep = ""),
                                             batterid = playerID)
    scrape5 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-29", sep = ""),
                                             end_date, batterid = playerID)
    statcastScrape <- rbind(scrape1, scrape2, scrape3, scrape4, scrape5)
  }
  if (month == 10) {
    statcastScrape <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-01", sep = ""), end_date, batterid =  playerID)
    
  }
  return(statcastScrape)
}
```

```{r}
##scrapes 1 month worth of statcast data use if statements to help generalize format since assuming not inputting correct format
scrapeStatcastmonth <- function(year, month, playerID){
  if (month < 10) {
    scrape1 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-01", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-07", sep = ""),
                                             batterid = playerID)
    scrape2 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-08", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-14", sep = ""),
                                             batterid = playerID)
    scrape3 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-15", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-21", sep = ""),
                                             batterid = playerID)
    scrape4 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-22", sep = ""),
                                             paste(as.character(year), "-0", as.character(month), "-28", sep = ""),
                                             batterid = playerID)
    if (month %in% c(3, 5, 7, 8)) {
      scrape5 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-29", sep = ""),
                                               paste(as.character(year), "-0", as.character(month), "-31", sep = ""),
                                               batterid = playerID)
    }
    else if (month %in% c(4, 6, 9)) {
      scrape5 <- scrape_statcast_savant_batter(paste(as.character(year), "-0", as.character(month), "-29", sep = ""),
                                               paste(as.character(year), "-0", as.character(month), "-30", sep = ""),
                                               batterid = playerID)
    }
  }
  else if (month == 10) {
    scrape1 <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-01", sep = ""),
                                             paste(as.character(year), "-", as.character(month), "-07", sep = ""),
                                             batterid = playerID)
    scrape2 <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-08", sep = ""),
                                             paste(as.character(year), "-", as.character(month), "-14", sep = ""),
                                             batterid = playerID)
    scrape3 <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-15", sep = ""),
                                             paste(as.character(year), "-", as.character(month), "-21", sep = ""),
                                             batterid = playerID)
    scrape4 <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-22", sep = ""),
                                             paste(as.character(year), "-", as.character(month), "-28", sep = ""),
                                             batterid = playerID)
    scrape5 <- scrape_statcast_savant_batter(paste(as.character(year), "-", as.character(month), "-29", sep = ""),
                                             paste(as.character(year), "-", as.character(month), "-30", sep = ""),
                                             batterid = playerID)
  }
  statcastScrape <- rbind(scrape1, scrape2, scrape3, scrape4, scrape5)
  return(statcastScrape)
}
```

```{r}
#scrapes full year by calling month function in loop
scrapeStatcastyear <- function(year, month, year_start, year_end, playerID){
  for (i in 1:length(month)) {
    Sys.sleep(5)
    if (i == 1) {
      yearScrape <- scrapeStatcastFirstmonth(year, month[i], year_start, playerID)
    }
    else if (i == length(month)) {
      holderScrape <- scrapeStatcastLastmonth(year, month[i], year_end, playerID)
      yearScrape %>% bind_rows(holderScrape) -> yearScrape
    }
    else {
      holderScrape <- scrapeStatcastmonth(year, month[i], playerID)
      yearScrape %>% bind_rows(holderScrape) -> yearScrape
    }
  }
  return(yearScrape)
}
```

```{r}
##months of each mlb season
months2015 <- c(4:10)
months2016 <- c(4:10)
months2017 <- c(4:10)
months2018 <- c(3:10)
months2019 <- c(3:9)
months2021 <- c(4:10)
months2022 <- c(4:10)
months2023 <- c(3:10)
months2024 <- c(3:9)
```

```{r}
judge <- 592450
ohtani <- 660271
acuna <- 660670
goldy <- 502671
harper <- 547180
trout <- 545361
belly <- 641355
betts <- 605141
yelich <- 592885
altuve <- 514888
stanton <- 519317
bryant <- 592178
dondalson <- 518626
```

```{r}
##scrapes each individual year with MVP winner as player
donaldson2015 <- scrapeStatcastyear(2015, months2015, "2015-04-05", "2015-10-04", dondalson)
harper2015 <- scrapeStatcastyear(2015, months2015, "2015-04-05", "2015-10-04", harper)

trout2016 <- scrapeStatcastyear(2016, months2016, "2016-04-03", "2016-10-01", trout)
bryant2016 <- scrapeStatcastyear(2016, months2016, "2016-04-03", "2016-10-01", bryant)

altuve2017 <- scrapeStatcastyear(2017, months2017, "2017-04-02", "2017-10-01", altuve)
stanton2017 <- scrapeStatcastyear(2017, months2017, "2017-04-02", "2017-10-01", stanton)

betts2018 <- scrapeStatcastyear(2018, months2018, "2018-03-29", "2018-10-01", betts)
yelich2018 <- scrapeStatcastyear(2018, months2018, "2018-03-29", "2018-10-01", yelich)

trout2019 <- scrapeStatcastyear(2019, months2019, "2019-03-20", "2019-09-29", trout)
belly2019 <- scrapeStatcastyear(2019, months2019, "2019-03-20", "2019-09-29", belly)

ohtani2021 <- scrapeStatcastyear(2021, months2021, "2021-04-01", "2021-10-03", ohtani)
harper2021 <- scrapeStatcastyear(2021, months2021, "2021-04-01", "2021-10-03", harper)

judge2022 <- scrapeStatcastyear(2022, months2022, "2022-04-07", "2022-10-05", judge)
goldy2022 <- scrapeStatcastyear(2022, months2022, "2022-04-07", "2022-10-05", goldy)

ohtani2023 <- scrapeStatcastyear(2023, months2023, "2023-03-30", "2023-10-01", ohtani)
acuna2023 <- scrapeStatcastyear(2023, months2023, "2023-03-30", "2023-10-01", acuna)

judge2024 <- scrapeStatcastyear(2024, months2024, "2024-03-28", "2024-09-30", judge)
ohtani2024 <- scrapeStatcastyear(2024, months2024, "2024-03-28", "2024-09-30", ohtani)
```

```{r}
##re filter by batter
donaldson2015 %>%
  filter(batter == 518626) -> donaldson2015
harper2015 %>%
  filter(batter == 547180) -> harper2015
season2015 <- rbind(donaldson2015, harper2015)

trout2016 %>%
  filter(batter == 545361) -> trout2016
bryant2016 %>%
  filter(batter == 592178) -> bryant2016
season2016 <- rbind(trout2016, bryant2016)

altuve2017 %>%
  filter(batter == 514888) -> altuve2017
stanton2017 %>%
  filter(batter == 519317) -> stanton2017
season2017 <- rbind(altuve2017, stanton2017)

betts2018 %>%
  filter(batter == 605141) -> betts2018
yelich2018 %>%
  filter(batter == 592885) -> yelich2018
season2018 <- rbind(betts2018, yelich2018)

trout2019 %>%
  filter(batter == 545361) -> trout2019
belly2019 %>%
  filter(batter == 641355) -> belly2019
season2019 <- rbind(trout2019, belly2019)

ohtani2021 %>%
  filter(batter == 660271) -> ohtani2021
harper2021 %>%
  filter(batter == 547180) -> harper2021
season2021 <- rbind(ohtani2021, harper2021)

judge2022 %>%
  filter(batter == 592450) -> judge2022
goldy2022 %>%
  filter(batter == 502671) -> goldy2022
season2022 <- rbind(judge2022, goldy2022)

ohtani2023 %>%
  filter(batter == 660271) -> ohtani2023
acuna2023 %>%
  filter(batter == 660670) -> acuna2023
season2023 <- rbind(ohtani2023, acuna2023)

judge2024 %>%
  filter(batter == 592450) -> judge2024
ohtani2024 %>%
  filter(batter == 660271) -> ohtani2024
season2024 <- rbind(judge2024, ohtani2024)
```

```{r}
mvpWinners <- rbind(season2015, season2016, season2017, season2018, season2019,
                    season2021, season2022, season2023, season2024)
```

```{r}
write_csv(mvpWinners, "C:/Users/Andrew Fish/Documents/Syracuse/SAL 602/Module 8/mvpWinners.csv")
```

The next section of the code will be data manipulation and ShinyApp
```{r}
mvpWinners <- read_csv("mvpWinners.csv")
masterID <- read_csv("masterid.csv")

##want to have a column for exit velo now only have x,y,z components
##formula taken from fangraphs to find exit velo from components
mvpWinners %>%
  mutate(x2 = vx0 ^ 2,
         y2 = vy0 ^ 2,
         z2 = vz0 ^ 2,
         exit_velo = sqrt(x2 + y2 + z2) * 0.6818) -> mvpWinners

##possible batters for drop down menu
mvpWinners %>%
  group_by(batter) %>%
  summarise() %>%
  inner_join(masterID, by = c("batter" = "mlb_id")) %>%
  select(batter, mlb_name) %>%
  arrange(mlb_name) -> possibleBatters

possibleBattersList <- possibleBatters$batter
names(possibleBattersList) <- possibleBatters$mlb_name

##creating strike zone
plate_width <- 17 + (9 / pi)
k_zone_plot <- ggplot(NULL, aes(x = plate_x, plate_z)) +
  geom_rect(xmin = -(plate_width / 2) / 12,
            xmax = (plate_width / 2) / 12,
            ymin = 1.5,
            ymax = 3.6, color = "black", alpha = 0) +
  coord_equal() +
  scale_x_continuous("Horizontal Location (ft.)",
                     limits = c(-2, 2)) +
  scale_y_continuous("Vertical Location (ft.)",
                     limits = c(0, 5))


ui <- fluidPage(
  titlePanel("Exit Velocity of MVP Winners - 2015-2024 (Excluding 2020)"),
  sidebarLayout(
    sidebarPanel(position = "left",
                  selectInput(inputId = "selectBatter",
                              label = h3("Batter"),
                              choices = possibleBattersList,
                              selected = 660271)
    ),
    sidebarPanel(position = "left",
      checkboxGroupInput(inputId = "years",
                         label = "Year",
                         choices = list("2015: Donaldson, Harper" = 2015,
                                        "2016: Trout, Bryant" = 2016,
                                        "2017: Altuve, Stanton" = 2017,
                                        "2018: Betts, Yelich" = 2018,
                                        "2019: Trout, Bellinger" = 2019,
                                        "2021: Ohtani, Harper" = 2021,
                                        "2022: Judge, Goldschmidt" = 2022,
                                        "2023: Ohtani, Acuña" = 2023,
                                        "2024: Judge, Ohtani" = 2024),
                         selected = 2024),
      helpText("To get image to show must have year selected to the year the player won MVP.
               You can have multiple years selected if player won multiple MVPs during timeframe.
               To help MVP Winners are labeled with each year.")
    )
  ),
  mainPanel(position = "right",
    plotOutput(outputId = "batterEV"),
    plotOutput(outputId = "exitveloHist")
  )
)


server <- function(input, output) {
  SelectedBatterReactive <- reactive({
    mvpWinners %>% filter(batter == input$selectBatter)
  })
  
  output$batterEV <- renderPlot({
    SelectedBatterReactive() %>%
      filter(game_year %in% input$years) -> selectedYears
    
    zones <- selectedYears %>%
      group_by(zone) %>%
      summarize(
        N = n(),
        right_edge = min(1.5, max(plate_x)),
        left_edge = max(-1.5, min(plate_x)),
        top_edge = min(5, quantile(plate_z, 0.95, na.rm = TRUE)),
        bottom_edge = max(0, quantile(plate_z, 0.05, na.rm = TRUE)),
        avg_ev = sum(exit_velo) / n(),
        plate_x = mean(plate_x),
        plate_z = mean(plate_z))
    
    k_zone_plot %+% zones +
      geom_rect(aes(xmax = right_edge, xmin = left_edge,
                    ymax = top_edge, ymin = bottom_edge,
                    fill = avg_ev, alpha = avg_ev),
                color = "darkgreen") +
      geom_text_repel(size = 4, aes(label = round(avg_ev, 2),
                                    color = avg_ev < 0.5)) +
      scale_fill_gradient(low = "blue", high = "red") +
      scale_color_manual(values = c("white", "black")) +
      guides(color = FALSE, alpha = FALSE) +
      ggtitle("Exit Velocity by Zone")
  })
  
  output$exitveloHist <- renderPlot({
    SelectedBatterReactive() %>%
      filter(game_year %in% input$years) -> selectedYears

    ggplot(selectedYears, aes(x = exit_velo, color = pitch_type)) +
      geom_histogram() + ggtitle("Exit Velocity Distribution") +
      xlab("Exit Velocity (MPH)") + ylab("n")
  })
}

shinyApp(ui = ui, server = server)
```

Bennet BABIP

```{r}
MVPStatcastList <- read_csv("MVPStatcastEra.csv")
FranchiseRosters <- read_csv("FranchiseRosters.csv")
```

```{r}
view(MVPStatcastList)
```

## We will be taking a look at the Batting Average on Balls in Play for MVP Winners, and the rosters they had during their MVP Seasons. 

## BABIP is important because it can distinguish the "luck" a player had throughout the season, typically the BABIP for players is higher than their normal average.

```{r}
# Franchise Rosters EXCLUDING MVP Winner of that Season (Mutating for Batting Average on Balls in Play)
FranchiseRosters %>% group_by(Year,Team, Player, BA, AB, H, SO, SF, WAR,HR, `Player-additional`) %>%
  select(Year, Team, Player, BA, AB, H, SO, SF, WAR, HR, `Player-additional`) %>%
  summarize(BABIP = (H-HR)/(AB-SO-HR+SF)) -> FranchiseBABIP
```

```{r}
## Scatterplot of BABIP of all Batters on the Franchises who had an MVP Winner (Statcast Era)
FranchiseBABIP %>% plot_ly(x = ~BA, y = ~BABIP, color = ~Team, text = ~Year) %>% add_markers()
```


```{r}
FranchiseBABIP %>% filter(Team != "Los Angeles Angels") %>% filter(Team != "Los Angeles Dodgers") %>% filter(Team != "New York Yankees") -> SoloMVP
ggplot(SoloMVP, aes(x = Team, y = BABIP, fill = Team)) +
  geom_boxplot() + coord_flip() + scale_fill_ordinal()
```

```{r}
## Histogram of Amount of Players who had a specific OPS based on each Franchise (Statcast Era MVP's Only)
ggplot(FranchiseBABIP, aes(x = BABIP, fill = Team)) +
  geom_histogram(breaks = seq(0.2, 1.0, by = 0.1)) +
  scale_fill_ordinal()
```
```{r}
ggplot(FranchiseRosters, aes(x = Team, y = WAR, fill = Team)) +
  geom_boxplot() + coord_flip() + scale_fill_ordinal()
```
```{r}
## On-Base Percentage and On-Base + Slugging Percentage by each Franchise Roster in the Statcast Era
library(plotly)
FranchiseRosters %>% plot_ly(x = ~OBP, y = ~OPS, color = ~Team, text = ~Year) %>% add_markers()
```

```{r}
## WAR and Games Played for the MVP Winners in the Statcast Era
library(plotly)
MVPStatcastList %>% plot_ly(x = ~WAR, y = ~G, color = ~Team, text = ~`Player-additional`) %>% add_markers()
```

```{r}
# Mutating for BABIP for the MVP Winners in the Statcast Era
MVPStatcastList %>% group_by(Year,Team, Player, BA, AB, H, SO, SF, WAR,HR, `Player-additional`) %>%
  select(Year, Team, Player, BA, AB, H, SO, SF, WAR, HR, `Player-additional`) %>%
  summarize(BABIP = (H-HR)/(AB-SO-HR+SF)) -> MVPBABIP
View(MVPBABIP)
```
```{r}
## Scatterplot of MVP Winners Batting Average and BABIP in their MVP Season
library(plotly)
MVPBABIP %>% plot_ly(x = ~BABIP, y = ~BA, color = ~`Player-additional`, text = ~Year) %>% add_markers(text = ~`Player-additional`)
```

```{r}
library(plotly)
MVPBABIP %>% plot_ly(x = ~BABIP) %>% add_boxplot() -> BABIPBoxPlot
BABIPBoxPlot
```

Xander Swing Rates

```{r}
mvpWinners %>%
  filter(description != "blocked_ball") %>%
  mutate(swing = case_when(
    type == "S" & description == "called_strike" ~ FALSE,
    type == "X" | type == "S" ~ TRUE,
    type == "B" ~ FALSE))-> mvpWinners
```

```{r}
plate_width <- 17 + (9 / pi)
k_zone_plot <- ggplot(NULL, aes(x = plate_x, plate_z)) +
  geom_rect(xmin = -(plate_width / 2) / 12,
            xmax = (plate_width / 2) / 12,
            ymin = 1.5,
            ymax = 3.6, color = "black", alpha = 0) +
  coord_equal() +
  scale_x_continuous("Horizontal Location (ft.)",
                     limits = c(-2, 2)) +
  scale_y_continuous("Vertical Location (ft.)",
                     limits = c(0, 5))
```

```{r}
ui <- fluidPage(
  titlePanel("MVP Swing Rates"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput(inputId = "selectPlayer",
                  label = h3("Select Player"),
                  choices = unique(mvpWinners$player_name),
                  selected = NULL),
      
      checkboxGroupInput(inputId = "types",
                         label = h3("Types of Outcomes"),
                         choices = list("S" = "S",
                                        "B" = "B",
                                        "X" = "X"),
                         selected = c("S", "B", "X"))
    ),
    mainPanel(
      plotOutput(outputId = "swingRateChart"),
      plotOutput(outputId = "launchSpeedAnglePlot")
    )
  )
)

server <- function(input, output) {
  
  output$swingRateChart <- renderPlot({
    req(input$selectPlayer)
    
    mvpWinners %>%
      filter(player_name == input$selectPlayer) %>%
      filter(type %in% input$types) -> SelectedPitchResults
    
    zones <- SelectedPitchResults %>%
      group_by(zone) %>%
      summarize(
        N = n(),
        right_edge = min(1.5, max(plate_x)),
        left_edge = max(-1.5, min(plate_x)),
        top_edge = min(5, quantile(plate_z, 0.95, na.rm = TRUE)),
        bottom_edge = max(0, quantile(plate_z, 0.05, na.rm = TRUE)),
        swing_pct = sum(swing == TRUE) / n(),
        plate_x = mean(plate_x),
        plate_z = mean(plate_z))
    
    k_zone_plot %+% zones +
      geom_rect(aes(xmax = right_edge, xmin = left_edge,
                    ymax = top_edge, ymin = bottom_edge,
                    fill = swing_pct, alpha = swing_pct),
                color = "darkgreen") +
      geom_text_repel(size = 3, aes(label = round(swing_pct, 2),
                                    color = swing_pct < 0.5)) +
      scale_fill_gradient(low = "blue", high = "red") +
      scale_color_manual(values = c("white", "black")) +
      guides(color = FALSE, alpha = FALSE) +
      ggtitle(paste("Swing Rates for", input$selectPlayer))
  })
  
  output$launchSpeedAnglePlot <- renderPlot({
    req(input$selectPlayer)
    
    player_data <- mvpWinners %>%
      filter(player_name == input$selectPlayer)
    
    ggplot(player_data, aes(x = launch_speed, y = launch_angle, color = cut(launch_speed, breaks = c(-Inf, 90, 100, Inf), labels = c("Blue", "Gray", "Red")))) +
      geom_point(alpha = 0.5) +
      scale_color_manual(values = c("Blue" = "blue", "Gray" = "gray", "Red" = "red")) +
      labs(title = paste("Launch Speed vs. Launch Angle for", input$selectPlayer),
           color = "Launch Speed") +
      xlab("Launch Speed") +
      ylab("Launch Angle")
  })
  
}

shinyApp(ui = ui, server = server)


```

Derek Voting Correlation
# Creating a Data Frame With Necessary Player Statistics from MVP Season
```{r}
# Note: All Stats Collected from Baseball Reference

PLAYER <- c("Aaron Judge, 2024", "Shohei Ohtani, 2024", "Shohei Ohtani, 2023", "Ronald Acuna, 2023", "Aaron Judge, 2022", "Paul Goldschmidt, 2022", "Shohei Ohtani, 2021", "Bryce Harper, 2021", "Mike Trout, 2019", "Cody Bellinger, 2019", "Mookie Betts, 2018", "Christian Yelich, 2018", "Jose Altuve, 2017", "Giancarlo Stanton, 2017", "Mike Trout, 2016", "Kris Bryant, 2016", "Josh Donaldson, 2015", "Bryce Harper, 2015")

RBI <- c(144, 130, 95, 106, 131, 115, 100, 84, 104, 115, 80, 110, 81, 132, 100, 102, 123, 99)

R <- c(122, 134, 102, 149, 133, 106, 103, 101, 110, 121, 129, 118, 112, 123, 123, 121, 122, 118)

TmR <- c(815, 842, 739, 947, 807, 772, 723, 734, 769, 886, 876, 754, 896, 778, 717, 808, 891, 703)

HR <- c(58, 54, 44, 41, 62, 35, 46, 35, 45, 47, 32, 36, 24, 59, 29, 39, 41, 42)

HITS <- c(180, 197, 151, 217, 177, 178, 138, 151, 137, 170, 180, 187, 204, 168, 173, 176, 184, 172)

TB <- c(392, 411, 325, 383, 391, 324, 318, 300, 303, 351, 333, 343, 323, 377, 302, 334, 352, 338)

BB <- c(133, 81, 91, 80, 111, 79, 96, 100, 110, 95, 81, 68, 58, 85, 116, 75, 73, 124)

AB <- c(559, 636, 497, 643, 570, 561, 537, 488, 470, 558, 520, 574, 590, 597, 549, 603, 620, 521)

AVG <- c(.322, .310, .304, .337, .311, .317, .257, .309, .291, .305, .346, .326, .346, .281, .315, .292, .297, .330)

OBP <- c(.458, .390, .412, .416, .425, .404, .372, .429, .438, .406, .438, .402, .410, .376, .441, .385, .371, .460)

OPS <- c(1.159, 1.036, 1.066, 1.012, 1.111, .981, .965, 1.044, 1.083, 1.035, 1.078, 1.000, .957, 1.007, .991, .939, .939, 1.109)

WAR <- c(10.8, 9.2, 6.0, 8.2, 10.5, 7.7, 4.9, 5.9, 7.9, 8.6, 10.7, 7.3, 7.7, 7.9, 10.5, 7.3, 7.1, 9.7)

VOTES <- c(30, 30, 30, 30, 28, 22, 30, 17, 17, 19, 28, 29, 27, 10, 19, 29, 23, 30)

MVPStats <- data.frame(PlayerNameandYear = PLAYER, RBI = RBI, R = R, TeamRuns = TmR, HR = HR, H = HITS, TB = TB, BB = BB, AB = AB, BAVG = AVG, OBP = OBP, OPS = OPS, WAR = WAR, MVPFirstPlaceVotes = VOTES)
```

# Adding the Bill James Run Created Formula to the Data Frame 
```{r}
MVPStats <- MVPStats %>% mutate(JamesRC = (((H + BB) * TB)/ (AB + BB)))
```

# Adding a Runs Accounted For Column and Percentage of Team Runs Accounted For Column
```{r}
# Ensure that Runs and RBIs due to Home Runs are not Double Counted
MVPStats <- MVPStats %>% mutate(RunsAccountedFor = (R - HR) + RBI)

# Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(PercentTmRunsAccountedFor = RunsAccountedFor/TeamRuns)
```

# Creating a Scatter Plot to Analyze the Correlation Between the Percentage of Team Runs Accounted for and 1st Place MVP Votes 
```{r}
# Calculating the Pearson Correlation r
TmRunsAccountedCorrelation <- cor(MVPStats$PercentTmRunsAccountedFor, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2TmRunsAcct <- MVPStats %>% arrange(desc(PercentTmRunsAccountedFor)) %>% head(2)
Bottom2TmRunsAcct <- MVPStats %>% arrange(desc(PercentTmRunsAccountedFor)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankAccountedTmRuns = case_when(
  PlayerNameandYear %in% Top2TmRunsAcct$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2TmRunsAcct$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = PercentTmRunsAccountedFor, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankAccountedTmRuns)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 0.2125, y = 12,
           label = paste("r-value =", round(TmRunsAccountedCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("Percentage of Team Runs Accounted For") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between % of Team Runs Accounted for and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2TmRunsAcct, color = "black", nudge_y = 1, aes(label = paste(PlayerNameandYear, round(PercentTmRunsAccountedFor, 2), MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2TmRunsAcct, color = "black", aes(label = paste(PlayerNameandYear, round(PercentTmRunsAccountedFor, 2), MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```

# Creating a Scatter Plot to Analyze the Correlation Between Runs Accounted for and 1st Place MVP Votes
```{r}
# Calculating the Pearson Correlation r
RunsAccountedCorrelation <- cor(MVPStats$RunsAccountedFor, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2RunsAcct <- MVPStats %>% arrange(desc(RunsAccountedFor)) %>% head(2)
Bottom2RunsAcct <- MVPStats %>% arrange(desc(RunsAccountedFor)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankAccountedRuns = case_when(
  PlayerNameandYear %in% Top2RunsAcct$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2RunsAcct$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = RunsAccountedFor, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankAccountedRuns)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 180, y = 12,
           label = paste("r-value =", round(RunsAccountedCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("Runs Accounted For") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between Runs Accounted for and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2RunsAcct, color = "black", nudge_y = 3, aes(label = paste(PlayerNameandYear, RunsAccountedFor, MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2RunsAcct, color = "black", nudge_y = 2, aes(label = paste(PlayerNameandYear, RunsAccountedFor, MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```

# Creating a Scatter Plot to Analyze the Correlation Between James' Runs Created and 1st Place MVP Votes
```{r}
# Calculating the Pearson Correlation r
JamesRCCorrelation <- cor(MVPStats$JamesRC, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2JamesRC <- MVPStats %>% arrange(desc(JamesRC)) %>% head(2)
Bottom2JamesRC <- MVPStats %>% arrange(desc(JamesRC)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankJamesRC = case_when(
  PlayerNameandYear %in% Top2JamesRC$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2JamesRC$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = JamesRC, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankJamesRC)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 165, y = 14,
           label = paste("r-value =", round(JamesRCCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("Runs Created According to James' Equation") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between Runs Created and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2JamesRC, color = "black", nudge_y = 3, aes(label = paste(PlayerNameandYear, round(JamesRC, 0), MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2JamesRC, color = "black", nudge_y = 1, aes(label = paste(PlayerNameandYear, round(JamesRC, 0), MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```


# Creating a Scatter Plot to Analyze the Correlation Between Player OPS and 1st Place MVP Votes
```{r}
# Calculating the Pearson Correlation r
OPSCorrelation <- cor(MVPStats$OPS, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2OPS <- MVPStats %>% arrange(desc(OPS)) %>% head(2)
Bottom2OPS <- MVPStats %>% arrange(desc(OPS)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankOPS = case_when(
  PlayerNameandYear %in% Top2OPS$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2OPS$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = OPS, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankOPS)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 1.1, y = 14,
           label = paste("r-value =", round(OPSCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("On-Base Plus Slugging Percentage") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between OPS and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2OPS, color = "black", nudge_y = -1, aes(label = paste(PlayerNameandYear, round(OPS, 2), MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2OPS, color = "black", nudge_y = -0.125, aes(label = paste(PlayerNameandYear, round(OPS, 2), MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```

# Creating a Scatter Plot to Analyze the Correlation Between WAR and 1st Place MVP Votes
```{r}
# Calculating the Pearson Correlation r
WARCorrelation <- cor(MVPStats$WAR, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2WAR <- MVPStats %>% arrange(desc(WAR)) %>% head(2)
Bottom2WAR <- MVPStats %>% arrange(desc(WAR)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankWAR = case_when(
  PlayerNameandYear %in% Top2WAR$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2WAR$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = WAR, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankWAR)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 9.5, y = 14,
           label = paste("r-value =", round(WARCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("Wins Above Replacement") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between WAR and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2WAR, color = "black", nudge_y = -1, aes(label = paste(PlayerNameandYear, round(WAR, 1), MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2WAR, color = "black", nudge_y = 1, aes(label = paste(PlayerNameandYear, round(WAR, 1), MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```

# Creating a Scatter Plot to Analyze the Correlation Between Home Runs and 1st Place MVP Votes
```{r}
# Calculating the Pearson Correlation r
HRCorrelation <- cor(MVPStats$HR, MVPStats$MVPFirstPlaceVotes)

# Determining Which Players are the Top and Bottom 2 in Terms of Percentage of Team Runs Accounted For 
Top2HR <- MVPStats %>% arrange(desc(HR)) %>% head(2)
Bottom2HR <- MVPStats %>% arrange(desc(HR)) %>% tail(2)

# Adding A new Column to MVP Stats to Assign Groups Based on Top or Bottom 2 of Percentage of Team Runs Accounted For 
MVPStats <- MVPStats %>% mutate(RankHR = case_when(
  PlayerNameandYear %in% Top2HR$PlayerNameandYear ~ "Top 2",
  PlayerNameandYear %in% Bottom2HR$PlayerNameandYear ~ "Bottom 2",
  TRUE ~ "Other"))

ggplot(MVPStats, aes(x = HR, y = MVPFirstPlaceVotes)) +
  geom_point(aes(color = RankHR)) +
  
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  
  annotate("text", x = 35, y = 14,
           label = paste("r-value =", round(HRCorrelation, 2)),
           size = 6, color = "black") +
  
  xlab("Home Runs") + ylab("Number of First Place MLB Votes") +
  
  ggtitle("Correlation Between Home Runs and 1st Place MVP Votes") +
  
  geom_text_repel(data = Top2HR, color = "black", aes(label = paste(PlayerNameandYear, HR, MVPFirstPlaceVotes))) +
  
  geom_text_repel(data = Bottom2HR, color = "black", aes(label = paste(PlayerNameandYear, HR, MVPFirstPlaceVotes))) +
  
  scale_color_manual(values = c("Top 2" = "green", "Bottom 2" = "red", "Other" = "black"))
```

# Conclusion of Analysis
```{r}
# Bill James' Runs Created Equation has the strongest correlation with MVP First Place voting. As a player creates more runs according to James, their MVP votes increase. There are other slightly positive correlations for other areas such as OPS, WAR, and Runs Accounted For. Surprisingly, there is a slightly negative correlation when looking at Home Runs and Percentage of Team Runs Accounted For compared to First Place MVP Votes. The voting committee does not seem to value how the MVP affects their team's success. 
```


# Writing New Data Frame With Stat Rankings to CSV 
```{r}
write_csv(MVPStats, "FinalProjectMVPStats.csv")
```
